---
title: "How to use the 'make_art' function"
output: html_document
date: "2023-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define the function

```{r}
make_art <- function(seed      = 42,
                     n_lines   = 500,
                     n_steps   = 80,
                     step_size = 200,
                     step_slip = 70,
                     spread    = 2,
                     prob_dist = "uniform",
                     generator = "simplex",
                     palette   = "davos",
                     animate   = FALSE) {
  
  # Load required packages
  require(tidyverse)
  require(ambient)
  require(scico)
  require(gganimate)
  require(randomcoloR)
  
  # List of parameters
  art_params <- list(seed      = seed,
                     n_lines   = n_lines,
                     n_steps   = n_steps,
                     step_size = step_size,
                     step_slip = step_slip,
                     spread    = spread,
                     prob_dist = prob_dist,
                     generator = generator,
                     palette   = palette)
  
  
  # Canvas setup -----------------------------------------------
  
  set.seed(art_params$seed)
  
  # Uniform distribution
  if (prob_dist == "uniform") {
    # Create tibble with initial coordinates for x and y. The z variable is
    # used to manipulate `step_slip` later
    state <- tibble(x = runif(n   = art_params$n_lines, 
                              min = 0, 
                              max = art_params$spread),
                    y = runif(n   = art_params$n_lines, 
                              min = 0, 
                              max = art_params$spread),
                    z = 0)
    
    # Add IDs per line and state (so we can track each line and state easily)
    state <- state %>% 
      mutate(line_id = 1:art_params$n_lines,
             step_id = 1)
    
    # This object will track the series of states per line
    art_data <- state
  }
  
  # Normal distribution
  else if (prob_dist == "normal") {
    # Create tibble with initial coordinates for x and y. The z variable is
    # used to manipulate `step_slip` later
    state <- tibble(x = rnorm(n    = art_params$n_lines, 
                              mean = 1, 
                              sd   = art_params$spread),
                    y = rnorm(n    = art_params$n_lines, 
                              mean = 1, 
                              sd   = art_params$spread),
                    z = 0)
    
    # Add IDs per line and state (so we can track each line and state easily)
    state <- state %>% 
      mutate(line_id = 1:art_params$n_lines,
             step_id = 1)
    
    # This object will track the series of states per line
    art_data <- state  
  }
  
  
  # Generator loop -----------------------------------------------
  
  # A while-loop that mutates the state until `n_steps` is reached
  stop_draw <- FALSE
  while (stop_draw == FALSE) {
    
    # Choosing different noise generating functions
    # Simplex
    if (generator == "simplex") {
      # Generate curl noise and add to coordinates
      step <- curl_noise(generator = gen_simplex,
                         x         = state$x,
                         y         = state$y,
                         z         = state$z,
                         seed      = rep(1, 3) * art_params$seed)
    }
    
    # Waves
    else if (generator == "waves") {
      # Generate curl noise and add to coordinates
      step <- curl_noise(generator = gen_waves,
                         x         = state$x,
                         y         = state$y,
                         z         = state$z,
                         seed      = rep(1, 3) * art_params$seed)
    }
    
    # Perlin
    else if (generator == "perlin") {
      # Generate curl noise and add to coordinates
      step <- curl_noise(generator = gen_perlin,
                         x         = state$x,
                         y         = state$y,
                         z         = state$z,
                         seed      = rep(1, 3) * art_params$seed)
    }
    
    # Spheres
    else if (generator == "spheres") {
      # Generate curl noise and add to coordinates
      step <- curl_noise(generator = gen_spheres,
                         x         = state$x,
                         y         = state$y,
                         z         = state$z,
                         seed      = rep(1, 3) * art_params$seed)
    }
    
    # Incrementally increase coordinates by small amount
    state <- state %>% 
      mutate(x       = x + (step$x / 10000) * art_params$step_size, 
             y       = y + (step$y / 10000) * art_params$step_size,
             z       = z + (step$z / 10000) * art_params$step_slip,
             step_id = step_id + 1)         
    
    # Append new states to `art_data`
    art_data <- bind_rows(art_data, state)
    
    # Halts while-loop once `n_steps` is reached
    if (last(state$step_id) >= art_params$n_steps) {
      stop_draw <- TRUE
    }
  }
  
  
  # Generate image and present -----------------------------------
  
  # Object for animation
  total_step_id <- 1:length(art_data$x)
  times <- data.frame(total_step_id)
  
  # Set color palette and randomly sample background color
  user_palette <- art_params$palette
  all_bgs      <- scico(n = 8, palette = user_palette)
  our_bg       <- sample(all_bgs, 1)
  
  art <- ggplot(data    = art_data,
                mapping = aes(x     = x, 
                              y     = y, 
                              color = step_id,
                              group = line_id)) +
    geom_path(linewidth   = 0.6, 
              alpha       = 0.5,
              show.legend = FALSE) + 
    coord_equal() +
    theme_void() +
    scale_color_scico(palette = user_palette) +
    theme(panel.background = element_rect(fill = our_bg))
  
  # Output static art
  if (animate == FALSE) {
    return(art) #	( ☉ ‿ ⚆ )
  }
  
  # Output animated art
  else if (animate == TRUE) {
    anim <- art + transition_reveal(times$total_step_id)
    return(anim)
  }
  
}
```

## Examples of Static Art

### Easy (Should Take a Few Seconds)

Here is an example of a (static) picture that should take only a few seconds to generate. You can try messing around with the parameters listed below. The `seed` determines the random seed used to randomly generate lines, `n_lines` determines how many lines are drawn, `n_steps` determines how iterations the code runs for (more iteratrions means longer lines), `step_size` determines how much is added onto a line per iteration, `step_slip` adds randomness and depth to all lines, `spread` determines how spread our the initial points from which lines are drawn will be, `prob_dist` can be used to choose between randomly sampling from 'unifrom' distributions or 'normal' distributions, `generator` can be used to select the noise generating function (determines the overall shape of the lines, choose between: 'simplex', 'waves', 'perlin', or 'spheres'), `palette` determines the color palette (palettes include all those found in the `scico` package), and finally `animate` allows us to choose between a static image or animated image.

```{r, dpi = 200}
make_art(seed      = 111,
         n_lines   = 100,
         n_steps   = 80,
         step_size = 200,
         step_slip = 80,
         spread    = 4,
         prob_dist = "uniform",
         generator = "simplex",
         palette   = "lajolla",
         animate   = FALSE)
```

### Medium (Should Take A Minute or So)

```{r, dpi = 200}
make_art(seed      = 911,
         n_lines   = 250,
         n_steps   = 100,
         step_size = 200,
         step_slip = 80,
         spread    = 4,
         prob_dist = "uniform",
         generator = "simplex",
         palette   = "lapaz",
         animate   = FALSE)
```

### Hard (Will Take a While)


```{r, dpi = 200}
make_art(seed      = 420,
         n_lines   = 600,
         n_steps   = 100,
         step_size = 200,
         step_slip = 80,
         spread    = 4,
         prob_dist = "uniform",
         generator = "simplex",
         palette   = "bamO",
         animate   = FALSE)
```



